---
title: "DK Map slavevoyages"
author: "Frida Lützow Kirk"
date: "2026-05-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

options(repos = c(CRAN = "https://cloud.r-project.org"))

```

----
# Map of the showing the Danish Trans-Atlantic slave trade


First, we install the packages necessary for creating the map
```{r packages}

install.packages("leaflet")
install.packages("tidyverse")
##For this type of project, i will make a html - "htmltools" is the one working for this program
install.packages("htmltools")
install.packages("dplyr")
```

Then we have to unpack everything in the library 
```{r library}
library(leaflet)
library(tidyverse)
library(htmltools)
library(dplyr)

```


Here i define the csv of the data on the danish slave voyages
```{r data}
danish_slave_trade <- read_csv("taden.csv")

danish_slave_trade
```

To decide what coordinates to show on our map, we are looking at which ports were the most frequent for disimbarkment for the captives

```{r finalports}
finalports <- danish_slave_trade %>% 
  group_by(`Imputed principal port of captive disembarkation`) %>% 
  count()

finalports
```

The most traveled to port is None - None is NA
We will remove None from our list

```{r finalports_nonone}
finalports_no_none <- finalports %>% 
  filter(`Imputed principal port of captive disembarkation`!="None")

```

Additionally on the map, we want to see the starting point for the vessel, and the place where the enslaved were purchased
Repeating the processes from above

```{r vesselstarting_point}
vessel_starting_point <- danish_slave_trade %>% 
  group_by(`Imputed Place where vessel's voyage began`) %>% 
  count()

vessel_starting_point

vessel_starting_point_no_none <- vessel_starting_point %>% 
  filter(`Imputed Place where vessel's voyage began` !="None")

```

Now we want a overview over where most people were purchaset 
```{r purchaseplace}

purchase_place <- danish_slave_trade %>% 
  group_by(`Imputed principal place of captive purchase`) %>% 
  count()

purchase_place 

purchase_place_no_none <- purchase_place %>% 
  filter(`Imputed principal place of captive purchase`!="None")

purchase_place_no_none

```

We are interested in seeing the top 3 locations for each of the lists
this is visible in the counted lists

In a google sheet we have collected the cordinates for these 9 places for ploting them into a leflet 

libraries needed for the map with google sheet coordinators

```{r packages2}
install.packages("leaflet")

install.packages("googlesheets4")


```

```{r libraryunpack}

library(tidyverse)
library(googlesheets4)
library(leaflet)

```

First we are making the base map
Here we need to open the providers for the map, we use Esri

```{r basemap}
leaflet() %>% 
  addTiles() %>% 
  setView( lng = 2.34, lat = 48.85, zoom = 5 ) %>%  
  addProviderTiles("Esri.WorldPhysical", group = "Physical") %>%  
  addProviderTiles("Esri.WorldImagery", group = "Aerial") %>%  
  addProviderTiles("MtbMap", group = "Geo") %>%               
  
  addLayersControl(                                 
    baseGroups = c("Geo","Aerial", "Physical"),
    options = layersControlOptions(collapsed = F)) 

esri <- grep("^Esri", providers, value = TRUE)
esri

base_map <- leaflet() %>% 
  setView(-20.770465174792065,19.10732583182444, zoom = 3) 
for (provider in esri) {
  base_map <- base_map %>% addProviderTiles(provider, group = provider)
}
base_map


dk_slave_voyages <- base_map %>%
  addProviderTiles("Esri.NatGeoWorldMap", group = "Aerial") %>% 
  addLayersControl(baseGroups = names(esri),
                   options = layersControlOptions(collapsed = TRUE)) %>%
  addMiniMap(tiles = esri[[1]], toggleDisplay = TRUE,
             position = "bottomright") %>%
  addMeasure(
    position = "bottomleft",
    primaryLengthUnit = "meters",
    primaryAreaUnit = "sqmeters",
    activeColor = "#3D535D",
    completedColor = "#7D4479") %>% 
  htmlwidgets::onRender("
                        function(el, x) {
                        var myMap = this;
                        myMap.on('baselayerchange',
                        function (e) {
                        myMap.minimap.changeLayer(L.tileLayer.provider(e.name));
                        })
                        }") %>% 
  addControl("", position = "topright")

dk_slave_voyages

library(htmlwidgets)

saveWidget(dk_slave_voyages, "baselayer_slavetrademap.html", selfcontained = TRUE)


```

Here we read in the Google sheet and select the Starting Port

In case there are difficulties with the read_sheet() function (it is erroring out), 
Run the following function:
gs4_deauth() 
And then rerun the read_sheet() function below

```{r deauth}
gs4_deauth()
```

```{r googleshet}

ship_starting_port <- read_sheet("https://docs.google.com/spreadsheets/d/1DrrjkuGPmcmtlis8zEXERD7Y7SZ-B0UxQHwLxZARgfE/edit?usp=sharing",
                     col_types = "cccnnc",   # we are working with 6 collumns c=is text n=numbers
                     range = "SP")  # here i have select the correct worksheet name

glimpse(ship_starting_port)

ship_starting_port %>% 
  filter(!is.na(Longitude)) %>% 
  filter(!is.na(Latitude))


```

Here we select the worksheet for the Captive Purchase

```{r googlesheet2}
captive_purchase_places <- read_sheet("https://docs.google.com/spreadsheets/d/1DrrjkuGPmcmtlis8zEXERD7Y7SZ-B0UxQHwLxZARgfE/edit?usp=sharing",
                                 col_types = "cccnnc",  
                                 range = "CP") 

glimpse(captive_purchase_places)

captive_purchase_places %>% 
  filter(!is.na(Longitude)) %>% 
  filter(!is.na(Latitude))

```

Here we select the worksheet for Captive Dismbarkment 

```{r googlesheet3}
captive_final_ports <- read_sheet("https://docs.google.com/spreadsheets/d/1DrrjkuGPmcmtlis8zEXERD7Y7SZ-B0UxQHwLxZARgfE/edit?usp=sharing",
                                 col_types = "cccnnc",  
                                 range = "CD")  

glimpse(captive_final_ports)

captive_final_ports %>% 
  filter(!is.na(Longitude)) %>% 
  filter(!is.na(Latitude))

```

# Final Map
Now we are testing to see if the markers show everything we want to show, before making the final map
Testing it with the CD worksheet

```{r finalmap}
test_cd<- leaflet() %>% 
  addTiles() %>% 
  addMarkers(lng = captive_final_ports$Longitude, 
             lat = captive_final_ports$Latitude,
             popup = paste(captive_final_ports$Placename,"<br>",captive_final_ports$Type,"<br>",captive_final_ports$Description))

test_cd




final_map <- dk_slave_voyages%>%
  addCircleMarkers(
    data = ship_starting_port,
    lng = ship_starting_port$Longitude, lat = ship_starting_port$Latitude,
    color = "blue",
    radius = 6,
    group = "ship_starting_port",
    popup = paste(ship_starting_port$Placename,"<br>",ship_starting_port$Type,"<br>",ship_starting_port$Description)
  ) %>%
  
  addCircleMarkers(
    data = captive_purchase_places,
    lng = captive_purchase_places$Longitude, lat = captive_purchase_places$Latitude,
    color = "red",
    radius = 6,
    group = "captive_purchase_places",
    popup = paste(captive_purchase_places$Placename,"<br>",captive_purchase_places$Type,"<br>",captive_purchase_places$Description)
  ) %>%
  
  addCircleMarkers(
    data = captive_final_ports,
    lng = captive_final_ports$Longitude, lat = captive_final_ports$Latitude,
    color = "hotpink",
    radius = 6,
    group = "captive_final_ports", 
    popup = paste(captive_final_ports$Placename,"<br>",captive_final_ports$Type,"<br>",captive_final_ports$Description)
  ) 
  

final_map

```

Now each dot on the map has a color
The blue dots represent ports where vessels started their voyages
The red dots represent Ports and Citys where enslaved people were bought
The green dots represent the final ports where the 


In the end, we want to save the html of the map

```{r map}
library(htmlwidgets)

saveWidget(final_map, "dkmap_slavetrade.html", selfcontained = TRUE)

```
Now we have a map showing the Trans Atlantic Slave trade from the Danish perspective


```{r pressure, echo=FALSE}

```
